---
title: "Untitled"
output: html_document
date: "2025-07-19"
---
```{r}
library("tidyverse", "rlang", "corrplot")
```

```{r}
testing <- read.csv("testing.csv")
```


Testing Imputation Quality
```{r}
testing %>% 
  select(acq_method, acq_method_filled) %>% 
  filter(acq_method == 'missing')

testing %>% 
  select(pol_edeliv_ind, pol_edeliv_ind_filled) %>% 
  filter(pol_edeliv_ind == -2)
```
















Call Count Analysis
```{r}
train <- read.csv("data/train_data.csv")
```


```{r}
summary(train$call_counts)
mean(train$call_counts)
var(train$call_counts)

sum(train$call_counts == 0) / length(train$call_counts)

```

```{r}
train %>%
  ggplot(aes(x=call_counts)) +
  geom_histogram(bins = 30, fill = "blue", color = "white") +
  labs(title = "Histogram of Call Counts",
       x = "Number of Calls",
       y = "Frequency") +
  theme_minimal()
```

Overdispersion exists in call_counts, which violates Poisson Distribution's assumption of mean and variance are both similar to lambda

Use Hurdle method to model call_counts:
https://stats.stackexchange.com/questions/81457/what-is-the-difference-between-zero-inflated-and-hurdle-models


Hurdle models assume that there is only one process by which a zero can be produced, while zero-inflated models assume that there are 2 different processes that can produce a zero.

Hurdle models assume 2 types of subjects: (1) those who never experience the outcome and (2) those who always experience the outcome at least once. Zero-inflated models conceptualize subjects as (1) those who never experience the outcome and (2) those who can experience the outcome but don't always.

In simple terms: both zero-inflated and hurdle models are described in two parts.

The first is the on-off part, which is a binary process. The system is "off" with probability π
 and "on" with probability 1−π
. (Here, π
 is known as the inflation probability.) When the system is "off," only zero counts are possible. This part is the same for zero-inflated and hurdle models.

The second part is the counting part, which occurs when the system is "on." This is where zero-inflated and hurdle models differ. In zero-inflated models, counts can still be zero. In hurdle models they must be nonzero. For this part, zero-inflated models use a "usual" discrete probability distribution while hurdle models use a zero-truncated discrete probability distribution function.


```{r}
# Check distribution for non-zero data

count_data <- train$call_counts[train$call_counts != 0]

summary(count_data)
mean(count_data)
var(count_data)


train %>%
  filter(train$call_counts != 0) %>% 
  ggplot(aes(x=count_data)) +
  geom_histogram(bins = 30, fill = "blue", color = "white") +
  labs(title = "Histogram of Call Counts",
       x = "Number of Calls",
       y = "Frequency") +
  theme_minimal()
```
Summary Finding:
- Try Zero-Inflated NegBin (handles overdispersion, excess zeros)
- Try Hurdle Model (first classification on (zero or nonzero) data then regression on count data)














Check step1 data

```{r}
X_train <- read.csv("data/model_data_auto/X_train.csv")
X_val <- read.csv("data/model_data_auto/X_val.csv")
X_test <- read.csv("data/model_data_auto/X_test.csv")

y_train <- read.csv("data/model_data_auto/y_train.csv")
y_val <- read.csv("data/model_data_auto/y_val.csv")

# X_train <- read.csv("data/model_data_Nonauto/X_train.csv")
# X_val <- read.csv("data/model_data_Nonauto/X_val.csv")
# X_test <- read.csv("data/model_data_Nonauto/X_test.csv")
# 
# y_train <- read.csv("data/model_data_Nonauto/y_train.csv")
# y_val <- read.csv("data/model_data_Nonauto/y_val.csv")

train <- cbind(X_train, y_train)
val <- cbind(X_val, y_val)

mean(train$call_counts)
var(train$call_counts)
```

Univariate Analysis

Numerical analysis using histogram
```{r}
histogram_summary <- function(data, num_var, bins = 30) {
  # Convert the variable name from string to symbol
  num_sym <- sym(num_var)

  # Histogram
  p <- data %>%
    ggplot(aes(x = !!num_sym)) +
    geom_histogram(bins = bins, fill = "steelblue", color = "black") +
    labs(
      title = paste("Histogram of", num_var),
      x = num_var,
      y = "Count"
    ) +
    theme_minimal()

  print(p)

  # Summary statistics
  summary_table <- data %>%
    summarize(
      mean = mean(!!num_sym, na.rm = TRUE),
      median = median(!!num_sym, na.rm = TRUE),
      sd = sd(!!num_sym, na.rm = TRUE),
      min = min(!!num_sym, na.rm = TRUE),
      max = max(!!num_sym, na.rm = TRUE),
      n = sum(!is.na(!!num_sym))
    )

  return(summary_table)
}
```

```{r}
histogram_summary(train, "ann_prm_amt")
histogram_summary(train, "call_counts")
# histogram_summary(train, "newest_veh_age") # only in auto
histogram_summary(train, "X12m_call_history")
histogram_summary(train, "tenure_at_snapshot")
histogram_summary(train, "trm_len_mo") # only in nonauto
histogram_summary(train, "home_lot_sq_footage")
histogram_summary(train, "household_policy_counts")
```

Nominal categorical analysis using boxplot

```{r}
categorical_summary <- function(data, cat_var) {
  # Convert variable name from string to symbol
  cat_sym <- sym(cat_var)

  # Bar plot
  p <- data %>%
    ggplot(aes(x = !!cat_sym)) +
    geom_bar(fill = "steelblue", color = "black") +
    labs(
      title = paste("Bar Plot of", cat_var),
      x = cat_var,
      y = "Count"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  print(p)

  total_n <- nrow(data)
  
  # Frequency table with proportions
  summary_table <- data %>%
    group_by(!!cat_sym) %>%
    summarize(
      count = n(),
      proportion = count / total_n,
      .groups = "drop"
    ) %>%
    arrange(desc(count))

  return(summary_table)
}

```


```{r}
categorical_summary(train, "acq_method_filled")
# categorical_summary(train, "bi_limit_group") # only in auto
categorical_summary(train, "channel") # only in non auto
categorical_summary(train, "digital_contact_ind")
categorical_summary(train, "geo_group")
categorical_summary(train, "has_prior_carrier")
categorical_summary(train, "household_group")
categorical_summary(train, "pay_type_code")
categorical_summary(train, "pol_edeliv_ind_filled")
categorical_summary(train, "prdct_sbtyp_grp")
categorical_summary(train, "product_sbtyp")
# categorical_summary(train, "telematics_ind") # only in auto
```




Bivariate Analysis
```{r}

boxplot_summary <- function(data, cat_var, num_var) {
  # Convert column names from strings to symbols
  cat_sym <- sym(cat_var)
  num_sym <- sym(num_var)

  # Boxplot
  p <- data %>%
    ggplot(aes(x = !!cat_sym, y = !!num_sym)) +
    geom_boxplot(fill = "skyblue", color = "darkblue", outlier.shape = 1, outlier.colour = "red") +
    labs(
      title = paste(num_var, "by", cat_var),
      x = cat_var,
      y = num_var
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  print(p)

  # Summary table
  summary_table <- data %>%
    group_by(!!cat_sym) %>%
    summarize(
      mean_value = mean(!!num_sym, na.rm = TRUE),
      median_value = median(!!num_sym, na.rm = TRUE),
      n = n(),
      .groups = "drop"
    ) %>%
    arrange(median_value)

  return(summary_table)
}

```



```{r}
# Is bi_limit_group truly nominal categorical?
boxplot_summary(train, "bi_limit_group", "ann_prm_amt")
boxplot_summary(train, "bi_limit_group", "call_counts")
boxplot_summary(train, "bi_limit_group", "newest_veh_age")
boxplot_summary(train, "bi_limit_group", "X12m_call_history")
boxplot_summary(train, "bi_limit_group", "tenure_at_snapshot")
boxplot_summary(train, "bi_limit_group", "home_lot_sq_footage")
boxplot_summary(train, "bi_limit_group", "household_policy_counts")

```

```{r}
# Is prdct_sbtyp_grp truly nominal categorical?
boxplot_summary(train, "prdct_sbtyp_grp", "ann_prm_amt")
boxplot_summary(train, "prdct_sbtyp_grp", "call_counts")
boxplot_summary(train, "prdct_sbtyp_grp", "newest_veh_age")
boxplot_summary(train, "prdct_sbtyp_grp", "X12m_call_history")
boxplot_summary(train, "prdct_sbtyp_grp", "tenure_at_snapshot")
boxplot_summary(train, "prdct_sbtyp_grp", "home_lot_sq_footage")
boxplot_summary(train, "prdct_sbtyp_grp", "household_policy_counts")
```





Correlation Matrix
```{r}
correlation_matrix <- function(data, num_vars) {

  # Select specified numeric columns
  selected_data <- data %>% select(num_vars)
  
  # Compute correlation matrix
  cor_matrix <- cor(selected_data, use = "pairwise.complete.obs")
  
  # Plot heatmap using corrplot
  corrplot::corrplot(cor_matrix,
                     method = "color",
                     type = "lower",
                     tl.col = "black",
                     addCoef.col = "black",  # Add correlation coefficients
                     number.cex = 0.7,
                     col = colorRampPalette(c("blue", "white", "orange"))(200),
                     title = "Correlation Matrix Heatmap",
                     mar = c(0, 0, 2, 0))  # Top margin space for title

  # Return correlation matrix
  return(cor_matrix)
}


```

```{r}
# Auto
library("corrplot")
correlation_matrix(train, c("call_counts", "ann_prm_amt", "newest_veh_age", "X12m_call_history", "tenure_at_snapshot", "home_lot_sq_footage", "household_policy_counts"))
```

```{r}
# Non Auto
library("corrplot")
correlation_matrix(train, c("call_counts", "ann_prm_amt",  "X12m_call_history", "tenure_at_snapshot", "home_lot_sq_footage", "household_policy_counts"))
```

